### 2019-09-24 final_NeurIPS/appendix.tex

%!TEX root = main.tex

\subsection{Proof of Lemma~\ref{lemma_actives_sets}}

Since the objective function $\Psi_{\kappa}$ is convex with respect to $(u,v)$, the set of optima of problem~\eqref{screen-sinkhorn} is non empty.
Introducing two dual variables $\lambda \in \R^n_{+}$ and $\beta \in \R^m_{+}$ for each constraint, the Lagrangian of problem~\eqref{screen-sinkhorn} reads as 
\begin{equation*}
  \mathscr{L}(u,v, \lambda, \beta) = \frac \varepsilon\kappa\inr{\lambda, \mathbf{1}_n} + \varepsilon\kappa\inr{\beta, \mathbf{1}_m} + \mathbf{1}_n^\top B(u,v) \mathbf{1}_m - \inr{\kappa u, \mu} - \inr{\frac v\kappa, \nu} -\inr{\lambda,e^{u}} - \inr{\beta,e^{v}}
\end{equation*}
First order conditions then yield that the Lagrangian multiplicators solutions $\lambda^{*}$ and $\beta^{*}$ satisfy 
\begin{align*}
  &\nabla_u\mathscr{L}(u^{*},v^{*}, \lambda^{*}, \beta^{*})=  e^{u^{*}} \odot(Ke^{v^{*}} - \lambda^{*}) - \kappa\mu = \mathbf 0_n,\\
  & \text{ and } \nabla_v\mathscr{L}(u^{*},v^{*}, \lambda^{*}, \beta^{*})=  e^{v^{*}} \odot(K^\top e^{u^{*}} - \beta) - \frac \nu\kappa = \mathbf 0_m
\end{align*}
which leads to 
\begin{align*}
  &\lambda^{*} = K e^{v^{*}} - \kappa\mu \oslash e^{u^{*}} \text{ and }
  \beta^{*} = K^\top e^{u^{*}} - \nu \oslash \kappa e^{v^{*}}
\end{align*}

For all $i=1, \ldots, n$ we have that $e^{u^{*}_i} \geq \frac\varepsilon\kappa$. Further, the condition on the dual variable $\lambda^{*}_i > 0$  ensures that $e^{u^{*}_i} = \frac\varepsilon\kappa$ and hence $i \in I^\complement_{\varepsilon,\kappa}$. We have that $\lambda^{*}_i > 0$ is equivalent to $e^{u^{*}_i}r_i(K) e^{v^{*}_j} >  \kappa{\mu_i}$ which  is satisfied when $\varepsilon^2r_i(K) >  \kappa{\mu_i}.$  
In a symmetric way we can prove the same statement for $e^{v^{*}_j}$.

\subsection{Proof of Proposition~\ref{prop:bounds_of_usc_and_vsc}}

We prove only the first statement~\eqref{bound_on_u} and similarly we can prove the second one~\eqref{bound_on_v}.
For all $i\in I_{\varepsilon,\kappa}$, we have $e^{u^{\text{sc}}_i} > \frac \varepsilon\kappa$ or $e^{u^{\text{sc}}_i} = \frac \varepsilon\kappa$. In one hand, if $e^{u^{\text{sc}}_i} > \frac \varepsilon\kappa$ then according to the optimality conditions $\lambda^{\text{sc}}_i = 0,$ which implies $e^{u^{\text{sc}}_i} \sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j} = \kappa\mu_i$.
In another hand, we have 
\begin{align*}
e^{u^{\text{sc}}_i} \min_{i,j}K_{ij} \sum_{j=1}^m e^{v^{\text{sc}}_j} \leq e^{u^{\text{sc}}_i} \sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j} = \kappa\mu_i.
\end{align*}
We further observe that $\sum_{j=1}^m e^{v^{\text{sc}}_j} = \sum_{j \in J_{\varepsilon,\kappa}} e^{v^{\text{sc}}_j} + \sum_{j \in J^\complement_{\varepsilon,\kappa}} e^{v^{\text{sc}}_j} \geq \varepsilon\kappa |J_{\varepsilon,\kappa}| + \varepsilon\kappa |J^\complement_{\varepsilon,\kappa}|=\varepsilon\kappa m.$ Then
\begin{equation*}
\max_{i\in I_{\varepsilon,\kappa}} e^{u^{\text{sc}}_i} \leq \frac \varepsilon\kappa \vee \frac{\max_{i\in I_{\varepsilon,\kappa}}\mu_i}{m\varepsilon K_{\min}}.
\end{equation*}
Analogously, one can obtain for all $j\in J_{\varepsilon,\kappa}$
\begin{equation}
\label{upper_bound_v_potential}
\max_{j\in J_{\varepsilon,\kappa}}e^{v^{\text{sc}}_j} \leq \varepsilon\kappa \vee \frac{\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon K_{\min}}.
\end{equation}

Now, since $K_{ij} \leq 1$, we have 
\begin{align*}
e^{u^{\text{sc}}_i} \sum_{j=1}^m e^{v^{\text{sc}}_j} \geq e^{u^{\text{sc}}_i} \sum_{j=1}^m K_{ij}e^{v^{\text{sc}}_j} = \kappa\mu_i.
\end{align*}
Using~\eqref{upper_bound_v_potential}, we get 
\begin{align*}
\sum_{j=1}^m e^{v^{\text{sc}}_j} &= \sum_{j \in J_{\varepsilon,\kappa}} e^{v^{\text{sc}}_j} + \sum_{j \in J^\complement_{\varepsilon,\kappa}} e^{v^{\text{sc}}_j}
\leq \varepsilon\kappa |J^\complement_{\varepsilon,\kappa}| + \varepsilon\kappa \vee \frac{\max_{j\in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon K_{\min}} |J_{\varepsilon,\kappa}|.
\end{align*}
Therefore,
\begin{align*}
\min_{i \in I_{\varepsilon,\kappa}} e^{u^{\text{sc}}_i}  \geq \frac \varepsilon\kappa \vee \frac{\kappa\min_{I_{\varepsilon,\kappa}}\mu_i}{\varepsilon\kappa (m-m_b) + \varepsilon\kappa \vee \frac{\max_{j\in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon K_{\min}} m_b}.
\end{align*}

\subsection{Proof of Proposition~\ref{proposition_error_in_marginals}}

We define the distance function $\varrho: \R_+ \times \R_+ \mapsto [0, \infty]$ by $\varrho(a,b) = b - a + a \log(\frac ab).$
While $\varrho$ is not a metric, it is easy to see that $\varrho$ is not nonnegative and satisfies $\varrho(a,b) =0$ iff $a=b$.
The violations are computed through the following function: 
\begin{equation*}
	d_{\varrho}(\gamma,\beta) = \sum_{i=1}^n \varrho(\gamma_i,\beta_i), \text{ for } \gamma, \beta \in \R^n_+.
\end{equation*}
Note that if $\gamma,\beta$ are two vectors of positive entries, $d_{\varrho}(\gamma,\beta)$ will return some measurement on how far they are from each other. The next Lemma is from~\cite{khalilabid2018} (see Lemma 7 herein).
\begin{lemma}
\label{lem:pinsker}
For any $\gamma, \beta \in \R^n_+$, the following generalized Pinsker inequality holds 
\begin{align*}
\norm{\gamma - \beta}_1 \leq \sqrt{7 (\norm{\gamma}_1\wedge \norm{\beta}_1)d_{\varrho}(\gamma,\beta)}.
\end{align*}
\end{lemma}

The optimality conditions for $({u}^{\text{sc}}, {v}^{\text{sc}})$ entails 
\begin{align}
\label{i-th-marginal-mu} 
{\mu}^{\text{sc}}_i  &= 
\begin{cases}
e^{u^{\text{sc}}_i} \sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}, \text{ if  }i \in I_{\varepsilon,\kappa},\\
\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}, \text{ if  }i \in I^\complement_{\varepsilon,\kappa}
\end{cases}
=\begin{cases}
\kappa \mu_i, \text{ if  }i \in I_{\varepsilon,\kappa},\\
\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}, \text{ if  }i \in I^\complement_{\varepsilon,\kappa},
\end{cases}
\end{align}
and 
\begin{align}
\label{i-th-marginal-nu}
{\nu}^{\text{sc}}_j  &= 
\begin{cases}
e^{v^{\text{sc}}_j} \sum_{i=1}^n K_{ij} e^{u^{\text{sc}}_i}, \text{ if  }j \in J_{\varepsilon,\kappa},\\
\varepsilon\kappa\sum_{i=1}^n K_{ij} e^{u^{\text{sc}}_i}, \text{ if  }j \in J^\complement_{\varepsilon,\kappa}
\end{cases}
=\begin{cases}
\frac{\nu_j}{\kappa}, \text{ if  }j \in J_{\varepsilon,\kappa},\\
\varepsilon\kappa\sum_{i=1}^n K_{ij} e^{u^{\text{sc}}_i}, \text{ if  }j \in J^\complement_{\varepsilon,\kappa}.
\end{cases}
\end{align}

By~\eqref{i-th-marginal-mu}, we have
\begin{align*}
d_\varrho({\mu} ,{\mu}^{\text{sc}}) &= \sum_{i=1}^n  {\mu}^{\text{sc}}_i - {\mu}_i + {\mu}_i  \log\Big(\frac{{\mu}_i}{{\mu}^{\text{sc}}_i }\Big)\\
&= \sum_{i\in I_{\varepsilon,\kappa}} (\kappa-1)\mu_i - \mu_i\log(\kappa) + \sum_{i\in I^\complement_{\varepsilon,\kappa}}\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j} - \mu_i + \mu_i \log\Big(\frac{\mu_i}{\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}}\Big)\\
&= \sum_{i\in I_{\varepsilon,\kappa}} (\kappa-\log(\kappa)-1)\mu_i  + \sum_{i\in I^\complement_{\varepsilon,\kappa}}\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j} - \mu_i + \mu_i \log\Big(\frac{\mu_i}{\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}}\Big).
\end{align*}
Now by~\eqref{bound_on_v}, we have in one hand 
\begin{align*}
\sum_{i\in I^\complement_{\varepsilon,\kappa}}\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}&= \sum_{i\in I^\complement_{\varepsilon,\kappa}}\frac \varepsilon\kappa \Big(\sum_{j\in J_{\varepsilon,\kappa}}K_{ij} e^{v^{\text{sc}}_j} + \varepsilon \kappa\sum_{j\in J^\complement_{\varepsilon,\kappa}}K_{ij}\Big)\\
&\leq \sum_{i\in I^\complement_{\varepsilon,\kappa}}\frac \varepsilon\kappa \Big(m_b \max_{i,j}K_{ij}\frac{\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon K_{\min}} + (m - m_b)\varepsilon\kappa\max_{i,j}K_{ij}\Big) \\
&\leq (n-n_b)\Big(\frac{m_b\max_{j} \nu_j}{n\kappa K_{\min}} + (m- m_b) \varepsilon^2\Big).
\end{align*}
On the other hand, we get
\begin{align*}
\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}&=\frac \varepsilon\kappa \Big(\sum_{j\in J_{\varepsilon,\kappa}}K_{ij} e^{v^{\text{sc}}_j} + \varepsilon \kappa\sum_{j\in J^\complement_{\varepsilon,\kappa}}K_{ij}\Big)\\
&\geq m_bK_{\min} \frac{m\varepsilon^2K_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\kappa((n-n_b)m\varepsilon^2K_{\min} + m\varepsilon^2K_{\min} + n_b\kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}\\
&\qquad +\varepsilon^2 (m- m_b) K_{\min}\\
&\geq \frac{mm_b\varepsilon^2(K_{\min})^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\kappa((n-n_b)m\varepsilon^2K_{\min}+ m\varepsilon^2K_{\min} + n_b\kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}\\
&\qquad +\varepsilon^2 (m- m_b) K_{\min}\\
&\geq \frac{mm_b\varepsilon^2K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\kappa((n-n_b)m\varepsilon^2K_{\min}+ m\varepsilon^2K_{\min} + n_b\kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}.
\end{align*}
Then 
\begin{align*}
\frac{1}{\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}} &\leq \frac{\kappa((n-n_b)m\varepsilon^2K_{\min}+ m\varepsilon^2K_{\min} + n_b\kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}{mm_b\varepsilon^2 K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\\
&\leq \frac{\kappa(n-n_b+ 1)}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{n_b\kappa^2\max_{i\in I_{\varepsilon,\kappa}}\mu_i}{mm_b\varepsilon^2K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}.
\end{align*}
It entails 
\begin{align*}
&\sum_{i\in I^\complement_{\varepsilon,\kappa}}\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j} - \mu_i + \mu_i \log\Big(\frac{\mu_i}{\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}}\Big)\\
&\qquad \leq (n-n_b)\bigg(\frac{m_b}{n\kappa K_{\min}} + (m- m_b) \varepsilon^2 - \min_{i}\mu_i\\
&\qquad \qquad + \max_{i}\mu_i\log\Big(\frac{\kappa(n-n_b+ 1)\max_{i}\mu_i}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{n_b\kappa^2(\max_{i}\mu_i)^2}{mm_b\varepsilon^2 K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\Big)
\bigg).
\end{align*}
Therefore
\begin{align*}
d_\varrho({\mu},{\mu}^{\text{sc}}) &\leq n_b c_{\kappa}\max_{i} \mu_i + (n-n_b)\bigg(\frac{m_b\max_{j}\nu_j}{n\kappa K_{\min}} + (m- m_b) \varepsilon^2 - \min_{i}\mu_i\\
&\qquad + \max_{i} \mu_i\log\Big(\frac{\kappa(n-n_b+ 1)\max_{i} \mu_i}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{n_b\kappa^2(\max_{i} \mu_i)^2}{mm_b\varepsilon^2 K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\Big).
\end{align*}
Finally, by Lemma~\ref{lem:pinsker} we obtain
\begin{align*}
\norm{{\mu} -{\mu}^{\text{sc}}}^2_1 \leq & n_bc_{\kappa}\max_{i} \mu_i + 7(n-n_b)\bigg(\frac{m_b\max_{j}\nu_j}{n\kappa K_{\min}} + (m- m_b) \varepsilon^2 - \min_{i}\mu_i\\
&+ \max_{i} \mu_i\log\Big(\frac{\kappa(n-n_b+ 1)\max_{i} \mu_i}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{n_b\kappa^2(\max_{i} \mu_i)^2}{mm_b\varepsilon^2K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\Big).
\end{align*}
Following the same lines as above, we also have
\begin{align*}
\norm{{\nu} -{\nu}^{\text{sc}}}^2_1 \leq & m_bc_{\frac 1\kappa}\max_{i} \mu_i + 7(m-m_b)\bigg(\frac{n_b\kappa\max_{i}\mu_i}{mK_{\min}} + (n- n_b) \varepsilon^2 - \min_{j}\nu_j\\
&+ \max_{j} \nu_j\log\Big(\frac{(m-m_b+ 1)\max_{j} \nu_j}{n_b\kappa K_{\min}\min_{i \in I_{\varepsilon,\kappa}}\mu_i} + \frac{m_b(\max_{j} \nu_j)^2}{nn_b\varepsilon^2\kappa^2K_{\min}^2\min_{i \in I_{\varepsilon,\kappa}}\mu_i}\Big).%\bigg\}^{1/2}
\end{align*}
To get the closed forms~\eqref{marginal-error-mu} and~\eqref{marginal-error-nu}, we used the follwoing facts:
\begin{remark}
\label{rem:orders_of_epsilonappa}
 We have 
$\varepsilon = \bigO((mnK_{\min}^2)^{-1/4})$, $\kappa= \bigO(\sqrt{m/(nc_{\mu\nu}K_{\min})}),$ $\kappa^{-1} = \bigO(\sqrt{n/(mK_{\min}c_{\mu\nu}})$, $(\kappa/\varepsilon)^2 = \bigO( m^{3/2} / \sqrt{nK_{\min}}(c_{\mu\nu})^{3/2})$, and $(\varepsilon\kappa)^{-2} = \bigO(n^{3/2}/\sqrt{mK_{\min}}c^{3/2}_{\mu\nu}).$
\end{remark}

\subsection{Proof of Proposition~\ref{prop:objective-error}}

We first define $\widetilde{K}$ a rearrangement of $K$ with respect to the active sets $I_{\varepsilon,\kappa}$ and $J_{\varepsilon,\kappa}$as follows:
\begin{equation*}
\widetilde{K} = 
\begin{bmatrix}
K_{(I_{\varepsilon,\kappa}, J_{\varepsilon,\kappa})} & K_{(I_{\varepsilon,\kappa}, J^\complement_{\varepsilon,\kappa})}\\
K_{(I^\complement_{\varepsilon,\kappa}, J_{\varepsilon,\kappa})} &K_{(I^\complement_{\varepsilon,\kappa}, J^\complement_{\varepsilon,\kappa})}
\end{bmatrix}.
\end{equation*}
Setting $\dt{\mu} = (\mu_{I_{\varepsilon,\kappa}}^\top, \mu_{I^\complement_{\varepsilon,\kappa}}^\top)^\top$, $\dt{\nu} = (\nu_{J_{\varepsilon,\kappa}}^\top, \nu_{J^\complement_{\varepsilon,\kappa}}^\top)^\top$ and for each vectors $u \in \R^n$ and $v\in \R^m$ we set $\dt{u} =(u_{I_{\varepsilon,\kappa}}^\top, u_{I^\complement_{\varepsilon,\kappa}}^\top)^\top  \text{ and } \dt{v} =(v_{J_{\varepsilon,\kappa}}^\top, v_{J^\complement_{\varepsilon,\kappa}}^\top)^\top.$
We then have 
\begin{equation*}
\Psi_{\varepsilon, \kappa} (u,v) = \mathbf 1_n^\top \widetilde{K}(\dt{u}, \dt{v}) \mathbf 1_m - \kappa \dt{\mu}^\top \dt{u} - \kappa^{-1} \dt{\nu}^\top \dt{v},
\end{equation*}
and 
\begin{equation*}
\Psi_{} (u,v) = \mathbf 1_n^\top \widetilde{K}(\dt{u}, \dt{v}) \mathbf 1_m - \dt{\mu}^\top \dt{u} - \dt{\nu}^\top \dt{v},
\end{equation*}
where
\begin{equation*}
  \widetilde{B}(\dt{u}, \dt{v}) = \Delta(e^{\dt{u}})\widetilde{K} \Delta(e^{\dt{v}}).
\end{equation*}
Let us consider the convex function
\begin{equation*}
	(\hat u, \hat v) \mapsto \inr{\mathbf 1_n, \widetilde{B}(\dt{\hat u}^{\text{}} ,\dt{\hat v}^{\text{}})\mathbf 1_m} - \inr{\kappa\dt{\hat u}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} - \inr{\kappa^{-1}\dt{\hat v}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n}.
\end{equation*}
Gradient inequality of any convex function g at point $x_o$ reads as $g(x_o) \geq g(x) + \inr{\nabla g(x), x_o - x}, \text{ for all } x \in \textbf{dom}(g).$
Applying the latter fact to the above function at point ($u^{\star}, v^{\star})$ we obtain
\begin{align*}
\inr{\mathbf 1_n, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} &- \inr{\kappa\dt{u}^{\text{sc}}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} - \inr{\kappa^{-1}\dt{v}^{\text{sc}}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n}\\
& - \big(\inr{\mathbf 1_n, \widetilde{B}(\dt{u}^{\star} ,\dt{v}^{\star})\mathbf 1_m}  
- \inr{\kappa\dt{u}^{\star{}}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} 
- \inr{\kappa^{-1}\dt{v}^\star, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n} \big)\\
&\qquad \qquad  \leq \inr{\dt{u}^{\text{sc}} - \dt{u}^\star, (1-\kappa) \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} 
+ \inr{\dt{v}^{\text{sc}} - \dt{v}^\star, (1-\kappa^{-1}) \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n}.
\end{align*}
Moreover,
\begin{align*}
\Psi_{\varepsilon, \kappa} (u^{\text{sc}} ,v^{\text{sc}}) -\Psi(u^\star, v^\star)
&= \inr{\mathbf 1_n, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} - \inr{\kappa\dt{u}^{\text{sc}}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} - \inr{\kappa^{-1}\dt{v}^{\text{sc}}, \widetilde{B}(\dt{u}^{\text{sc}}, \dt{v}^{\text{sc}})\mathbf 1_n^\top}\\
&\qquad - \big( 
\inr{\mathbf 1_n, \widetilde{B}(\dt{u}^{\star} ,\dt{v}^{\star})\mathbf 1_m}  
- \inr{\dt{u}^{\star{}}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} 
- \inr{\dt{v}^\star, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n} \big)\\
&\qquad + \inr{\kappa \dt{u}^{\text{sc}} - \dt{u}^{\star}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m - \dt{\mu}} + \inr{\kappa^{-1}\dt{v}^{\text{sc}} - \dt{v}^{\star}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n - \dt{\nu}}.
\end{align*}
Hence,
\begin{align*}
\Psi_{\varepsilon, \kappa} (u^{\text{sc}} ,v^{\text{sc}}) -\Psi(u^\star, v^\star)
+& \big( \inr{\mathbf 1_n, \widetilde{B}(\dt{u}^{\star} ,\dt{v}^{\star})\mathbf 1_m} \\
&- \inr{\dt{u}^{\star{}}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} 
- \inr{\dt{v}^\star, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n} \big)\\
& \qquad -\inr{\kappa \dt{u}^{\text{sc}} - \dt{u}^{\star}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m - \dt{\mu}} -\inr{\kappa^{-1}\dt{v}^{\text{sc}} - \dt{v}^{\star}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n - \dt{\nu}}\\
& \leq \inr{\dt{u}^{\text{sc}} - \dt{u}^\star, (1-\kappa) \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} 
+ \inr{\dt{v}^{\text{sc}} - \dt{v}^\star, (1-\kappa^{-1}) \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n}\\
&\qquad + \big(\inr{\mathbf 1_n, \widetilde{B}(\dt{u}^{\star} ,\dt{v}^{\star})\mathbf 1_m}  
- \inr{\kappa\dt{u}^{\star{}}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} 
- \inr{\kappa^{-1}\dt{v}^\star, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n} \big).
\end{align*}
Then,
\begin{align*}
\Psi_{\varepsilon, \kappa} (u^{\text{sc}} ,v^{\text{sc}}) -\Psi(u^\star, v^\star)
&\leq \inr{\dt{u}^{\text{sc}} - \dt{u}^\star, (1-\kappa) \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} 
+ \inr{\dt{v}^{\text{sc}} - \dt{v}^\star, (1-\kappa^{-1}) \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n}\\
&\qquad + \big(\inr{\mathbf 1_n, \widetilde{B}(\dt{u}^{\star} ,\dt{v}^{\star})\mathbf 1_m}  
- \inr{\kappa\dt{u}^{\star{}}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} 
- \inr{\kappa^{-1}\dt{v}^\star, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n} \big)
\\
&\qquad + \inr{\kappa \dt{u}^{\text{sc}} - \dt{u}^{\star}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m - \dt{\mu}} + \inr{\kappa^{-1}\dt{v}^{\text{sc}} - \dt{v}^{\star}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n - \dt{\nu}}\\
&\qquad - \big(\inr{\mathbf 1_n, \widetilde{B}(\dt{u}^{\star} ,\dt{v}^{\star})\mathbf 1_m}  
- \inr{\dt{u}^{\star{}}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} 
- \inr{\dt{v}^\star, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top \mathbf 1_n} \big),
\end{align*}
which yields 
\begin{align*}
\Psi_{\varepsilon, \kappa} (u^{\text{sc}} ,v^{\text{sc}}) -\Psi(u^\star, v^\star)
&\leq  \inr{\kappa \dt{u}^{\text{sc}} - \dt{u}^{\star}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m - \dt{\mu}} +\inr{\kappa^{-1}\dt{v}^{\text{sc}} - \dt{v}^{\star}, \widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n - \dt{\nu}}\\
&\qquad + (1- \kappa)\inr{\dt{u}^{\text{sc}} ,\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} + (1- \kappa^{-1})\inr{\dt{v}^{\text{sc}} ,\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n}.
\end{align*}
Applying Holder's inequality gives 
\begin{align*}
\Psi_{\varepsilon, \kappa}(u^{\text{sc}} ,v^{\text{sc}}) -\Psi(u^\star, v^\star)
&\leq \norm{\kappa \dt{u}^{\text{sc}} - \dt{u}^\star}_\infty \norm{\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m - \dt{\mu}}_1 + \norm{\kappa^{-1}\dt{v}^{\text{sc}} - \dt{v}^\star}_\infty \norm{\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n - \dt{\nu}}_1\\
&\qquad + (1- \kappa)\inr{\dt{u}^{\text{sc}} ,\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} + (1- \kappa^{-1})\inr{\dt{v}^{\text{sc}} ,\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n}\\
&\leq \big(\norm{\dt{u}^{\text{sc}} - \dt{u}^\star}_\infty + (1 - \kappa) \norm{\dt{u}^{\text{sc}}}_\infty\big)\norm{\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m - \dt{\mu}}_1\\
&\qquad + \big(\norm{\dt{v}^{\text{sc}} - \dt{v}^\star}_\infty + (\kappa^{-1} -1) \norm{\dt{v}^{\text{sc}}}_\infty\big) \norm{\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n - \dt{\nu}}_1\\
&\qquad + (1- \kappa)\inr{\dt{u}^{\text{sc}} ,\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m} + (\kappa^{-1}- 1)\inr{\dt{v}^{\text{sc}} ,\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n}
\end{align*}
where, in the last inequality, we use the facts that $\norm{\kappa\dt{u}^{\text{sc}} - \dt{u}^\star}_\infty \leq \norm{\dt{u}^{\text{sc}} - \dt{u}^\star}_\infty + (1 - \kappa) \norm{\dt{u}^{\text{sc}}}_\infty$ and $\norm{\kappa^{-1}\dt{v}^{\text{sc}} - \dt{v}^\star}_\infty \leq \norm{\dt{v}^{\text{sc}} - \dt{v}^\star}_\infty + (\kappa^{-1} -1) \norm{\dt{v}^{\text{sc}}}_\infty.$
Moreover, note that 
\begin{align*}
\begin{cases}
\norm{\dt{u}^{\text{sc}} - \dt{u}^\star}_\infty = \norm{{u}^{\text{sc}} - {u}^\star}_\infty,\\
\norm{\dt{v}^{\text{sc}} - \dt{v}^\star}_\infty = \norm{{v}^{\text{sc}} - {v}^\star}_\infty,
\end{cases}
\text{ and }
\begin{cases}
\norm{\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})\mathbf 1_m - \dt{\mu}}_1 = \norm{{B}({u}^{\text{sc}} ,{v}^{\text{sc}})\mathbf 1_m - {\mu}}_1 = \norm{\mu^{\text{sc}} - \mu}_1,\\
\norm{\widetilde{B}(\dt{u}^{\text{sc}} ,\dt{v}^{\text{sc}})^\top\mathbf 1_n - \dt{\nu}}_1 = \norm{{B}({u}^{\text{sc}} ,{v}^{\text{sc}})^\top\mathbf 1_n - {\nu}}_1 = \norm{\nu^{\text{sc}} - \nu}_1.
\end{cases}
\end{align*}
Then
\begin{align}
\Psi_{\varepsilon, \kappa}(u^{\text{sc}} ,v^{\text{sc}}) -\Psi(u^\star, v^\star) &
\leq \big(\norm{{u}^{\text{sc}} - {u}^\star}_\infty + (1- \kappa) \norm{{u}^{\text{sc}}}_\infty\big)\norm{\mu^{\text{sc}} - \mu}_1\nonumber\\
&\qquad + \big(\norm{{v}^{\text{sc}} - {v}^\star}_\infty + (\kappa^{-1} - 1) \norm{{v}^{\text{sc}}}_\infty\big) \norm{\nu^{\text{sc}} - \nu}_1\nonumber\\
&\qquad + (1- \kappa)\inr{{u}^{\text{sc}} ,{\mu}^{\text{sc}}} + (\kappa^{-1} -1)\inr{{v}^{\text{sc}} ,{\nu}^{\text{sc}}}\nonumber\\
&\leq \big(\norm{{u}^{\text{sc}} - {u}^\star}_\infty + (1- \kappa) \norm{{u}^{\text{sc}}}_\infty\big)\norm{\mu^{\text{sc}} - \mu}_1\nonumber\\
&\qquad + \big(\norm{{v}^{\text{sc}} - {v}^\star}_\infty + (\kappa^{-1} - 1) \norm{{v}^{\text{sc}}}_\infty\big) \norm{\nu^{\text{sc}} - \nu}_1 \label{priori_bounds_induction}\\
&\qquad + (1- \kappa)\norm{{u}^{\text{sc}}}_\infty\norm{{\mu}^{\text{sc}}}_1 + (1- \kappa^{-1})\norm{{v}^{\text{sc}}}_\infty\norm{{\nu}^{\text{sc}}}_1.\nonumber
\end{align}
Next, we bound the two terms $\norm{{u}^{\textrm{sc}} - {u}^\star}_\infty $ and $\norm{{v}^{\text{sc}} - {v}^\star}_\infty.$
% \paragraph{Bounds of}$\norm{{u}^{\textrm{sc}} - {u}^\star}_\infty $ and $\norm{{v}^{\text{sc}} - {v}^\star}_\infty $.
If $r \in I^\complement_{\varepsilon, \kappa}$, then we have 
\begin{align*}
|({u}^{\text{sc}})_{r} - {u}^\star_r| 
&= \bigg|\log\bigg(\frac{\sum_{j=1}^m K_{rj}e^{v^\star_j}}{\sum_{j=1}^m \frac{\kappa \mu_r}{m\varepsilon}}\bigg) \bigg|\\
& \overset{(\star)}{\leq} \bigg|\log\bigg(\max_{1\leq i \leq m} \frac{K_{rj}e^{v^\star_j}}{\frac{\kappa \mu_r}{m\varepsilon}}\bigg) \bigg|\\
& \leq \big|\max_{1\leq j \leq m} (v_j^\star - \log(\frac{\kappa \mu_r}{m\varepsilon})\big|\\
& \leq \norm{v^\star - \log(\frac{\kappa \mu_r}{m\varepsilon})}_\infty \\
& \leq \norm{v^\star - {v}^{\text{sc}}}_\infty + \log(\frac{m\varepsilon^2}{c_{\mu\nu}}).
\end{align*}
where the inequality $(\star)$ comes from the fact that $\frac{\sum_{j=1}^na_j}{\sum_{j=1}^nb_{j}} \leq \max_{1\leq j \leq n} \frac{a_j}{b_j}, \forall a_j, b_j>0.$
Now, if $r \in I_{\varepsilon, \kappa}$, we get 
\begin{align*}
|{u}^{\text{sc}}_{r} - {u}^\star_r|
= \bigg|\log\bigg(\frac{\kappa\sum_{j=1}^m K_{rj}e^{v^\star_j}}{\sum_{j=1}^m K_{rj}e^{(v^{\text{sc}})_j}}\bigg)\bigg|
\leq \bigg|\log\bigg(\frac{\sum_{j=1}^m K_{rj}e^{v^\star_j}}{\sum_{j=1}^m K_{rj}e^{(v^{\text{sc}})_j}}\bigg)\bigg|
\overset{(\star)}{\leq}\norm{v^{\text{sc}} - v^\star}_\infty.
\end{align*}
If $s \in J^\complement_{\varepsilon, \kappa}$ then 
\begin{align*}
|{v}^{\text{sc}}_{s} - {v}^\star_s| &= \bigg| \log(\varepsilon\kappa) - \log(\frac{\nu_s}{\sum_{i=1}^n K_{is}e^{u^\star_i}})\bigg|\\
& \leq \bigg|\log\bigg(\max_{1\leq i \leq n} \frac{K_{is}e^{u^\star_i}}{\frac{\nu_s}{n\kappa\varepsilon}}\bigg) \bigg|\\
& \overset{(\star)}{\leq} \big|\max_{1\leq i \leq n} (u_i^\star - \log(\frac{\nu_s}{n\kappa\varepsilon})\big|\\
& \leq \norm{u^\star - \log(\frac{\nu_s}{n\kappa\varepsilon})}_\infty \\
& \leq \norm{u^\star - u^{\text{sc}}}_\infty +  \log(\frac{n\varepsilon^2}{c_{\mu\nu}}).
\end{align*}
If $s \in J_{\varepsilon, \kappa}$ then 
\begin{align*}
|{v}^{\text{sc}}_{s} - {v}^\star_s| 
= \bigg|\log\bigg(\frac{\kappa\sum_{i=1}^m K_{ri}e^{u^\star_i}}{\sum_{i=1}^m K_{ri}e^{(u^{\text{sc}})_i}}\bigg)\bigg|
 \leq \bigg|\log\bigg(\frac{\kappa\sum_{i=1}^m K_{ri}e^{v^\star_j}}{\sum_{i=1}^m K_{ri}e^{(u^{\text{sc}})_i}}\bigg)\bigg|
\overset{(\star)}{\leq} \norm{u^{\text{sc}} - u^\star}_\infty.
\end{align*}
Therefore, we obtain the followoing bound:
\begin{align}
\label{induction_uppers_bounds}
\max\{\norm{u^\star - u^{\text{sc}}}_\infty, \norm{v^\star - v^{\text{sc}}}_\infty\} %&\leq \max\{\norm{u^\star - u^{\text{sc}}_0}_\infty +  \log(\frac{n\varepsilon^2}{c_{\mu\nu}}), \norm{v^\star - v^{\text{sc}}_0}_\infty + \log(\frac{m\varepsilon^2}{c_{\mu\nu}})\}\nonumber\\
&\leq \max\Big\{\norm{u^\star}_\infty + \norm{u^{\text{sc}}}_\infty +  \log(\frac{n\varepsilon^2}{c_{\mu\nu}}), \norm{v^\star}_\infty + \norm{v^{\text{sc}}}_\infty + \log(\frac{m\varepsilon^2}{c_{\mu\nu}})\Big\}\nonumber\\
&\leq 2\Big(\norm{u^\star}_\infty + \norm{v^\star}_\infty +  \norm{u^{\text{sc}}}_\infty +  \norm{v^{\text{sc}}}_\infty + \log\big(\frac{(n\vee m)\varepsilon^2}{c_{\mu\nu}}\big)\Big).
\end{align}
%where we used $u^{\text{sc}}_0 = \mathbf 0_n$ and $v^{\text{sc}}_0 = \mathbf 0_m.$ 
Now, Lemma 3.2 in~\cite{lin2019} provides an upper bound for the $\ell_\infty$ of the optimal solution pair $(u^\star, v^\star)$ of problem~\eqref{sinkhorn-dual} as follows:
$\norm{u^\star}_\infty \leq R$ and $\norm{v^\star}_\infty\leq R,$ where 
\begin{align}
\label{R_constant}
R = \frac{\norm{C}_\infty}{\eta} + \log\big(\frac{n\vee m}{c^2_{\mu\nu}}\big).
\end{align}
Plugging~\eqref{induction_uppers_bounds} and~\eqref{R_constant} in~\eqref{priori_bounds_induction}, we obtain
\begin{align}
\Psi_{\varepsilon, \kappa}(u^{\text{sc}} ,v^{\text{sc}}) -\Psi(u^\star, v^\star) 
&\leq 2\big(R + \norm{u^{\text{sc}}}_\infty +  \norm{v^{\text{sc}}}_\infty + \log\big(\frac{(n\vee m)\varepsilon^2}{c_{\mu\nu}})\big) \big(\norm{\mu^{\text{sc}} - \mu}_1+ \norm{\nu^{\text{sc}} - \nu}_1\big)\nonumber\\
&\qquad + (1- \kappa)\big(\norm{{u}^{\text{sc}}}_\infty\norm{{\mu}^{\text{sc}}}_1 +\norm{\mu^{\text{sc}} - \mu}_1\big)\label{post_bounds_induction_2}\\
&\qquad  + (\kappa^{-1} - 1)\big(\norm{{v}^{\text{sc}}}_\infty\norm{{\nu}^{\text{sc}}}_1+\norm{\nu^{\text{sc}} - \nu}_1\big)\nonumber.
\end{align}
By Proposition~\ref{prop:bounds_of_usc_and_vsc}, we have 
\begin{align*} %\max_{i \in I_{\varepsilon,\kappa}} \mu_i %\max_{j \in J_{\varepsilon,\kappa}} \nu_j
\norm{u^{\text{sc}}}_\infty \leq \log\big(\frac \varepsilon\kappa\vee \frac{1}{m\varepsilon K_{\min}}\big) \text{ and } \norm{v^{\text{sc}}}_\infty \leq \log\big(\varepsilon\kappa \vee \frac{1}{n\varepsilon K_{\min} }\big)
\end{align*}
and hence by Remark~\ref{rem:orders_of_epsilonappa}, 
\begin{equation*}
\norm{u^{\text{sc}}}_\infty = \bigO\big(\log({n^{1/4}}/{(mK_{\min})^{3/4}c_{\mu\nu}^{1/4}})\big) \text{ and }\norm{u^{\text{sc}}}_\infty =\bigO\big(\log({m^{1/4}}/{(nK_{\min})^{3/4}c_{\mu\nu}^{1/4}})\big).
\end{equation*}
Acknowledging that $\log(1/ K_{\min}^2) = 2\norm{C}_\infty/\eta$, we have 
\begin{align*}
R + \norm{u^{\text{sc}}}_\infty  + \norm{v^{\text{sc}}}_\infty + \log\big(\frac{(n\vee m)\varepsilon^2}{c_{\mu\nu}})\big) = \bigO\Big(\frac{\norm{C}_\infty}{\eta} + \log\Big(\frac{(n\vee m)^2}{nmc_{\mu\nu}^{7/2}}\Big)\Big).
\end{align*}
Letting $\Omega_{\kappa} := (1- \kappa)\big(\norm{{u}^{\text{sc}}}_\infty\norm{{\mu}^{\text{sc}}}_1 +\norm{\mu^{\text{sc}} - \mu}_1\big) + (\kappa^{-1} - 1)\big(\norm{{v}^{\text{sc}}}_\infty\norm{{\nu}^{\text{sc}}}_1+\norm{\nu^{\text{sc}} - \nu}_1\big).$
% Moreover, we have 
% \begin{align*}
% \omega_{\kappa} := (1- \kappa)\norm{{u}^{\text{sc}}}_\infty\norm{{\mu}^{\text{sc}}}_1 + (1- \kappa^{-1})\norm{{v}^{\text{sc}}}_\infty\norm{{\nu}^{\text{sc}}}_1
% \end{align*}
We have that 
\begin{align*}
\Omega_{\kappa} &=\bigO\Big(\Big(\frac{\norm{C}_\infty}{\eta} + \log\big(\frac{1}{(nm)^{3/4} c_{\mu\nu}^{1/2}}\big)\Big)\Big((1- \kappa)(\norm{{\mu}^{\text{sc}}}_1 +\norm{\mu^{\text{sc}} - \mu}_1\big) + (\kappa^{-1} - 1)(\norm{{\nu}^{\text{sc}}}_1+\norm{\nu^{\text{sc}} - \nu}_1)\Big)\Big)\\
&= \bigO\Big(\Big(\frac{\norm{C}_\infty}{\eta} + \log\Big(\frac{(n\vee m)^2}{nmc_{\mu\nu}^{7/2}}\Big)\Big)\big((1- \kappa)\norm{\mu^{\text{sc}}}_1 + (\kappa^{-1} - 1)\norm{\nu^{\text{sc}}}_1 + \kappa^{-1} - \kappa\big)\Big).
\end{align*}
% Setting $\omega_\kappa := (1- \kappa)\norm{\mu^{\text{sc}}}_1 + (\kappa^{-1} - 1)\norm{\nu^{\text{sc}}}_1 + \kappa^{-1} - \kappa,$ 
Hence, we arrive at 
\begin{align*}
\Psi_{\varepsilon, \kappa}(u^{\text{sc}} ,v^{\text{sc}}) -\Psi(u^\star, v^\star) 
& = \bigO\Big(\Big(\frac{\norm{C}_\infty}{\eta} + \log\Big(\frac{(n\vee m)^2}{nmc_{\mu\nu}^{7/2}}\Big)\Big)(\norm{\mu - {\mu}^{\text{sc}}}_1 + \norm{\nu - \nu^{\text{sc}}}_1 + \omega_{\kappa}\Big).
\end{align*}
$\hfill\square$

Note that $\omega_{\kappa}$ tends to zero as $(n_b,m_b)$ tends to $(n,m)$ (i.e., $\kappa \rightarrow 1)$. To more characterize $\omega_\kappa$, the following 
lemma expresses an upper bound with respect to $\ell_1$-norm of $\mu^{\text{sc}}$and $\nu^{\text{sc}}$. 
\begin{lemma}
\label{lemma_bounds_on_marginals}
Let $(u^{\text{sc}}, v^{\text{sc}})$ be an optimal solution of problem~\eqref{screen-sinkhorn_second_def}.
Then one has 
\begin{align}
\label{l1-norm-mu-sc}
\norm{\mu^{\text{sc}}}_1 = \bigO\Big(\frac{n_b\sqrt{m}}{\sqrt{nK_{\min}c_{\mu\nu}}} + (n-n_b)\Big(\frac{m_b}{\sqrt{nmc_{\mu\nu}}K_{\min}^{3/2}} + \frac{m-m_b}{\sqrt{nm}K_{\min}}\Big)\Big),
\end{align}
and
\begin{align}
\label{l1-norm-nu-sc}
\norm{\nu^{\text{sc}}}_1 = \bigO\Big(\frac{m_b\sqrt{n}}{\sqrt{mK_{\min}c_{\mu\nu}}} + (m-m_b)\Big(\frac{n_b}{\sqrt{nmc_{\mu\nu}}K_{\min}^{3/2}} + \frac{n-n_b}{\sqrt{nm}K_{\min}}\Big)\Big).
\end{align}
\end{lemma}

\begin{proof}
Using inequality~\eqref{bound_on_v}, we obtain 
\begin{align*}
\norm{\mu^{\text{sc}}}_1 &= \sum_{i \in I_{\varepsilon,\kappa}} \mu^{\text{sc}}_i +  \sum_{i \in I^\complement_{\varepsilon,\kappa}}\mu^{\text{sc}}_i\\
& \overset{\eqref{i-th-marginal-mu}}{=} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + \frac \varepsilon\kappa \sum_{i \in I^\complement}\Big( \sum_{j \in J_{\varepsilon,\kappa}} K_{ij} e^{v^{\text{sc}}_j} + \varepsilon\kappa \sum_{j\in J^\complement_{\varepsilon,\kappa}}K_{ij}\Big)\\
& \overset{\eqref{bound_on_v}}{\leq} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + (n-n_b) \Big(\frac{m_b\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\kappa K_{\min}} + (m-m_b)\varepsilon^2 \Big).
% &\leq n_b\kappa + (n-n_b) \Big(\frac{m_b\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\kappa K_{\min}} + (m-m_b)\varepsilon^2 \Big)
\end{align*}
Using Remark~\ref{rem:orders_of_epsilonappa}, we get the desired closed form in~\eqref{l1-norm-mu-sc}.
Similarly, we can prove the same statement for $\norm{\nu^{\text{sc}}}_1$.
\end{proof}

\ar{

\subsection{Additional remarks}

\paragraph{On the use of a constrained L-BFGS-B solver.}

It is worth to note that standard Sinkhorn's alternating minimization cannot be applied for the constrained screened dual problem~\eqref{screen-sinkhorn_second_def}. This appears more clearly while writing its optimality conditions (see Equations~\eqref{i-th-marginal-nu} and~\eqref{i-th-marginal-mu} ).
We resort to a L-BFGS-B algorithm to solve the constrained convex optimization problem on the screened variables~\eqref{screen-sinkhorn_second_def}, but any other efficient solver (e.g., proximal based method or Newton method) could be used. 
We further  emphasize that, as for the \textsc{Sinkhorn} algorithm, the \textsc{Screenkhorn} algorithm can be accelerated using a GPU implementation\footnote{\url{https://github.com/nepluno/lbfgsb-gpu}} of the L-BFGS-B algorithm~\cite{fei2014-GPU-L-BFGS-B}.

\paragraph{Comparison with other solvers.}

We have considered experiments with \textsc{Greenkhorn} algorithm~\cite{altschulernips17} but the implementation in POT library and our Python version of Matlab Alschuler's Greenkhorn code\footnote{\url{https://github.com/JasonAltschuler/OptimalTransportNIPS17}} were not competitive with \textsc{Sinkhorn}. Hence, for both versions, \textsc{Screenkhorn} is more competitive than \textsc{Greenkhorn}. The computation time gain reaches an order of $30$ when comparing our method with \textsc{Greenkhorn} while \textsc{Screenkhorn} is almost $2$ times faster than $\textsc{Sinkhorn}$.

\subsection{Additional experimental results}

\paragraph{Experimental setup.}

	All computations have been run on each single core of an Intel Xeon E5-2630 processor clocked at
	2.4 GHz in a Linux machine with 144 Gb of memory.
	The code that generates all figures given in the paper is available from \url{https://github.com/mzalaya/screenkhorn}.

}

\begin{figure*}[htbp]
	\centering
	%~\hfill~	\includegraphics[width=6.cm]{./figs/da_toy_problem.pdf}~\hfill~\\
	~\hfill~\includegraphics[width=6.cm]{./figs/wda_accur_toy.pdf}~\hfill~
	\includegraphics[width=6.cm]{./figs/wda_accur_mnist.pdf}~\hfill~\\
	\caption{Accuracy of a $1$-nearest-neighbour after WDA for the (left) toy problem and, (right) MNIST). We note a slight loss of performance for the toy problem, whereas for MNIST, all approaches yield the same performance. 
		\label{fig:wda_gain}}
\end{figure*}

\begin{figure*}[htbp]
	\centering
	%~\hfill~	\includegraphics[width=6.cm]{./figs/da_toy_problem.pdf}~\hfill~\\
	~\hfill~\includegraphics[width=6.cm]{./figs/da_toy_problem.pdf}~\hfill~
	\includegraphics[width=6.cm]{./figs/da_accur_toy_regcl100.pdf}~\hfill~\\
	~\hfill~\includegraphics[width=6.cm]{./figs/da_time_toy_regcl100.pdf}
	~\hfill~\includegraphics[width=6.cm]{./figs/da_gain_toy_regcl100.pdf}~\hfill~\\
	\caption{OT Domain Adaptation on a 3-class Gaussian toy problem. (top-left) Examples of source and target samples. (top-right) Evolution of the accuracy of a 1-nearest-neighbour classifier with respect to the number of samples. (bottom-left) Running time of the \textsc{Sinkhorn} and \textsc{Screenkhorn} for different decimation
	factors. (bottom-right). Gain in computation time. This toy problem is a problem in which classes are overlapping and distance between samples are rather limited. According to our analysis, this may be a situation in which \textsc{Screenkhorn} may
	result in smaller computational gain. We can remark that with respect to the accuracy
	\textsc{Screenkhorn} with decimation factors up to $10$ are competitive with \textsc{Sinkhorn}, although a slight loss of performance. Regarding computational time, for this example, small decimation factors does not  result in gain. However for above  $5$-factor decimation, the gain goes from  $2$ to $10$ depending on
	the number of samples. 
	\label{fig:otda:extra}}
\end{figure*}

\begin{figure*}[htbp]
	\centering
	\includegraphics[width=6.5cm]{./figs/da_accur_mnist_regcl1.pdf}
	\includegraphics[width=6.5cm]{./figs/da_accur_mnist_regcl10.pdf}
	\includegraphics[width=6.5cm]{./figs/da_time_mnist_regcl1.pdf}
	\includegraphics[width=6.5cm]{./figs/da_time_mnist_regcl10.pdf}
	\caption{OT Domain adaptation MNIST to USPS : (top) Accuracy and (bottom) running time of \textsc{Sinkhorn} and \textsc{Screenkhorn} for hyperparameter of the $\ell_{p,1}$ regularizer (left) $\lambda = 1$ and (right) $\lambda=10$. Note that this
	value impacts the ground cost of each Sinkhorn problem involved in the iterative algorithm. The accuracy panels
	also report the performance of a $1$-NN when no-adaptation is performed.  
	We remark that the strenght of the class-based regularization has influence on the performance of \textsc{Screenkhorn} given a decimation factor. For small value on the left, \textsc{Screenkhorn} slightly performs better than \textsc{Sinkhorn}, while for large value, some
	decimation factors leads to loss of performances.
	Regarding, running time, we can note that \textsc{Sinkhorn} is far less efficient than
	\textsc{Screenkhorn}  with an order of magnitude for intermediate number of samples.
	\label{fig:otda:mnist:extra}}
\end{figure*}